---
title: "R Independent Study"
author: "Enxhi Xhindoli"
date: "04/29/2019"
output:
  html_document: default
  pdf_document: default
---
In this report, a logistic regression was constructed in order to obtain the likelihood of an account churning based off variables utilized within the model. The logistic regression allows to assess the churn likelihood of pre-existing accounts.

```{r include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r include=FALSE, results="hide",warning=FALSE, message=FALSE}
library(rvest)
library(stringr)
library(tidyverse)
library(dplyr)
library(plyr)
library(here)
library(ROSE)
library(caret)
```


## Data Collection

The telecomunication company provided the LSU team with different datasets about demographics and rate changes. For the scope of this project, only a subset of the whole dataset was used. The variable BAN_SEQ is the ID which identifies a unique customer. 

```{r message=FALSE, include=FALSE }
RCU <- read.csv(here('Data', 'RateChangeUnique.csv'), sep =",", header = TRUE)
DEMO <- read.csv(here('Data', 'DEMO.csv'), sep =",", header=TRUE)
DF <- full_join(DEMO,RCU, by = "BAN_SEQ")
```

```{r include=FALSE}
write.csv(DF, file="C:/Users/enxhi/Desktop/R_independentstudy/R_project/Data/Rate_Demo.csv")
```

In order to add some information about the income, a table containing the average income per state and another table containing the states and their abbreviations were scraped from the website into the Global Environment of Rstudio. 

```{r income, include=FALSE}
income <- read_html("https://www.infoplease.com/business-finance/poverty-and-income/capita-personal-income-state")%>%
  html_node(css = "#A0104653")%>%
  html_table()
```

```{r include=FALSE}
#only the column showing the average income of each state in 2015 was kept
income <- as_tibble(income)
income <- select(income, State,`2015`)
income[1:3,]
```

```{r include=FALSE}
typeof(income)
```

```{r message=FALSE, include=FALSE}
income <- income %>%
  mutate(`2015`=str_extract_all(`2015`,"[0-9]+"))%>%
  mutate(`2015`=map_chr(`2015`,paste,collapse=""))

income <- as.data.frame(income)
names(income)[names(income) == "2015"] <- "AVG_INCOME_2015"
names(income)[names(income) == "State"] <- "STATE"
```


```{r state_abbreviation, include=FALSE}
#another table containing the states and their abbreviations was extracted from the web
states <- read_html("https://www.50states.com/abbreviations.htm") %>%
  html_node("[class='spaced stripedRows']")%>%
  html_table()

#Commonwealth territories were dropped from the dataset
states <- as.data.frame(states)
states <- states[-(51:67),]

names(states)[names(states) == "US State:"] <- "STATE"
names(states)[names(states) == "Abbreviation:"] <- "STATE_ABBREVIATION"
colnames(states)
```

```{r include=FALSE}
Income_per_state <- merge(states,income,by="STATE")
head(Income_per_state)
```

```{r warning=FALSE, include=FALSE}
#the two tables were merged with the csv dataset provided by the company
names(DF)[names(DF) == "STATE"] <- "STATE_ABBREVIATION"
DF1 <- left_join(DF,Income_per_state, by = "STATE_ABBREVIATION")
```

```{r include=FALSE}
write.csv(DF1, file="C:/Users/enxhi/Desktop/R_independentstudy/R_project/Data/Rate_Demo_Income.csv")
```

## Data preparation

In this analysis, only the subset of the observations belonging to the year 2015 will be considered.  

```{r results="hide", include=FALSE}
DF1$SNAPSHOT_DATE <- as.character(DF1$SNAPSHOT_DATE)
DF1$SNAPSHOT_DATE <- as.Date(DF1$SNAPSHOT_DATE, format = "%m/%d/%Y")
```

```{r message=FALSE, warning=FALSE, include=FALSE }
DF2 <-DF1 %>% 
  filter(SNAPSHOT_DATE>'2014-12-31' & SNAPSHOT_DATE<'2016-01-31')

length(unique(DF2$SNAPSHOT_DATE))
unique(DF2$SNAPSHOT_DATE)
```

From the dataset, it can be clearly seen that the variable CVGENDERCODE1, which is showing the gender, has more than two occurrences. For practical reasons, the name of the column was first changed to 'GENDER', and some inspections about it were made.

```{r include=FALSE}
names(DF2)[names(DF2) == "CVGENDERCODE1"] <- "GENDER"
DF2$GENDER <- as.character(DF2$GENDER)
```

```{r}
unique(DF2$GENDER)
```

The observations containing values such as: U and null were converted to NA. In this dataset, 22% of the obseravtions have missing values for gender. 

```{r results="hide"}
DF2$GENDER <- revalue(DF2$GENDER, c("U"=NA))
DF2$GENDER <- revalue(DF2$GENDER, c("null"=NA))
count(DF2$GENDER=='NA')
```

Inspections were made also for the variable EDUCATIONAL_LEVEL and the values '#N/D' were substituted with NA.

```{r include=FALSE}
unique(DF2$EDUCATION_LEVEL)
```

```{r}
DF2$EDUCATION_LEVEL <- revalue(DF2$EDUCATION_LEVEL, c("#N/D"=NA))
```

To see how the customer's paying rate has changed a new variable was created as difference between CURR_PAYING_RATE and PRIOR_PAYING_RATE.

```{r include=FALSE}
is.numeric(DF2$CURR_PAYING_RATE)
is.factor(DF2$PRIOR_PAYING_RATE)

DF2$CURR_PAYING_RATE <- as.character(DF2$CURR_PAYING_RATE)
DF2$CURR_PAYING_RATE <- as.numeric(DF2$CURR_PAYING_RATE)
DF2$CURR_PAYING_RATE[is.na(DF2$CURR_PAYING_RATE)]<- 0

DF2$PRIOR_PAYING_RATE <- as.character(DF2$PRIOR_PAYING_RATE)
DF2$PRIOR_PAYING_RATE <- as.numeric(DF2$PRIOR_PAYING_RATE)
DF2$PRIOR_PAYING_RATE[is.na(DF2$PRIOR_PAYING_RATE)]<- 0
```

```{r PAYING_RATE_CHANGE, echo=FALSE}
DF3 <- DF2 %>% mutate(PAYING_RATE_CHANGE= CURR_PAYING_RATE - PRIOR_PAYING_RATE)
```

Given the definition of customer churn as *"Customers who stopped using company's product or service during a certain time frame"*, for the creation of variable "CHURN" the CURR_QTY and PRIOR_QTY were taken into consideration. The two variables refer respectively to the current and prior quantity of products purchased by the single customer.

```{r}
DF3$CHURN <- ifelse(DF3$CURR_QTY >= DF3$PRIOR_QTY,0,1)
```

To make this analysis even more dynamic, the binary variable NEW_CUST was created. As the name reveals, it shows whether a person is a new customer in 2015 or not.

```{r}
DF3$NEW_CUST <- if_else(DF3$CURR_QTY > 0 & DF3$PRIOR_QTY == 0,1,0)
DF3$NEW_CUST <- factor(DF3$NEW_CUST)
```

```{r include=FALSE}
write.csv(DF3, file="C:/Users/enxhi/Desktop/R_independentstudy/R_project/Data/Final_Data.csv")
```

## Visualization

The first graph shows the trend of the number of customers and how many of those customers are "new customers" joining the company.

```{r results="hide", warning=FALSE, message=FALSE, include=FALSE}
detach(package:plyr, unload=TRUE)
library(ggplot2)
```

```{r echo=FALSE, message=FALSE}
V1 <- na.omit(DF3)
V2 <- V1%>% group_by(SNAPSHOT_DATE,NEW_CUST)%>% summarise(CUSTOMERS=n())
ggplot(V2, aes(SNAPSHOT_DATE,CUSTOMERS,fill=NEW_CUST)) + 
  xlab("Timeline") + ylab("Customers by Month") + geom_bar(stat = 'identity') +  theme(axis.text.x=element_text(angle=60, hjust=1))+ scale_x_date(breaks = V2$SNAPSHOT_DATE, date_labels = '%m/%y') 
```

To see in which state the company has the highest number of customers,a barchart was used. 

```{r echo=FALSE}
ggplot(V1,
       aes(x=reorder(STATE_ABBREVIATION,STATE_ABBREVIATION,
                     function(x)-length(x)))) + geom_bar()+ 
  theme(axis.text.x=element_text(angle=60, hjust=1)) + 
  xlab("State") + ylab("Number of customers")
```

PAYING_RATE_CHANGE is a variable created as the difference between the current paying rate and the prior paying rate by the customers.
To see for each credit class if the majority of the changes in paying rate has been negative or positive (respectively increasing and decreasing payments),a barchart graphic was created.

```{r echo=FALSE, message=FALSE, warning=FALSE}
theme_set(theme_bw())  
V1[V1$CREDIT_CLASS=="A"|V1$CREDIT_CLASS=="H"|V1$CREDIT_CLASS=="Z"|V1$CREDIT_CLASS=="U"|V1$CREDIT_CLASS=="P",] <- NA
V1$PAYING_RATE <- ifelse(V1$PAYING_RATE_CHANGE < 0, "decrease", "increase") 
V1 <- V1[order(V1$PAYING_RATE_CHANGE), ]  # sort
V1$CREDIT_CLASS <- factor(V1$CREDIT_CLASS)  # convert to factor to retain sorted order in plot.
V3 <- na.omit(V1)

# Diverging Barcharts
ggplot(V3, aes(x=CREDIT_CLASS, y=PAYING_RATE_CHANGE)) + 
  geom_bar(stat='identity', aes(fill=PAYING_RATE), width=.5)  +
  scale_fill_manual(name="Paying Rate", 
                    labels = c("Decreasing Paying Rate", "Increasing Paying Rate"), 
                    values = c("decrease"="#00ba38", "increase"="#f8766d")) + 
  labs(subtitle="Based on credit class", 
       title= "Change in paying rate") + 
  xlab("Credit class") + ylab("Change in paying rate")+
  coord_flip()
```

Furthermore, we want se see the correlation between the change in paying rate and the number of calls.

```{r echo=FALSE, message=FALSE, warning=FALSE}
V1$CALL_COUNT <- as.numeric(V1$CALL_COUNT)

ggplot(V1, aes(x = PAYING_RATE_CHANGE, y= CALL_COUNT)) + geom_point(color="red", size = 2, shape = 23) + xlab("Change in paying rate") + ylab("Calls")
```

As we can see from the graph, most of the calls are distributed around the interval [-50;50] of the PAYING_RATE_CHANGE, and most of the customers have a range of calls between 0 and 100. However, there are also some outliers as it can be seen from the graph above.

```{r message=FALSE, include=FALSE}
DF3$CREDIT_CLASS <- as.factor(DF3$CREDIT_CLASS)

DF3$EDUCATION_LEVEL <- factor(DF3$EDUCATION_LEVEL)
DF3$GENDER <- factor(DF3$GENDER)

DF3$AVG_INCOME_2015 <- as.numeric(DF3$AVG_INCOME_2015)

DF3$AGE <- as.character(DF3$AGE)
DF3$AGE <- as.numeric(DF3$AGE)
is.na(DF3$AGE) <- 0
```

## Builging a Logistic Regression model to predict CHURN

The objective of the Churn predictive model is to assign to the customers a measure of the propensity to abandonment (called score). However, before building a model, it is necessary to see how many instances of CHURN there are in the dataset.

```{r}
table(DF3$CHURN)
```

Since the event of churn is a rare event, to maximize the possibility of developing a model on a sample basis extensible in a robust way also on the total population, all the cases of abandonment detected in the analysis period were considered within the sample, while the cases of "no churn" were narrowed down (undersampling). The sample was therefore constructed as follows:  25% formed by units presenting the "abandonment" event (made up of all the customers lost during the analysis period) and 75% formed by units that do not have the "abandonment" event ( randomly extracted from all customers who have not left).

```{r message=FALSE, warning=FALSE}
DF4 <- na.omit(DF3)
# Split the data into training and test set
set.seed(123)
training.samples <- DF4$CHURN %>% 
  createDataPartition(p = 0.7, list = FALSE)
train.data  <- DF4[training.samples, ]
test.data <- DF4[-training.samples, ]
# imbalance on training set
table(train.data$CHURN)
# balanced data set with under-sampling
data.balanced.under <- ovun.sample(CHURN~CREDIT_CLASS+NEW_CUST+CALL_COUNT+AGE+GENDER+EDUCATION_LEVEL+AVG_INCOME_2015, data=train.data, p=0.25, seed=1, method="under")$data
table(data.balanced.under$CHURN)
```

Now a logistic regression model is created based on training balanced data.

```{r message=FALSE, warning=FALSE, echo=FALSE}
train_bal <- glm(CHURN~CREDIT_CLASS+NEW_CUST+CALL_COUNT+GENDER+EDUCATION_LEVEL+AVG_INCOME_2015, data=data.balanced.under, family=binomial)
summary(train_bal)
```

The model was then applied to the test data.
```{r message=FALSE, warning=FALSE, echo=FALSE}
# use the trained model to predict test data
probabilities<- predict(train_bal, newdata=test.data,
type="response")
# compute accuracy measures (test set)
accuracy.meas(test.data$CHURN, probabilities)
```

Finally, as it can be clearly seen from the results, this model is not at all accurated. This makes sense as the demographic data was collected from third parties, and therefore built upon many assumptions, but also because it is hard to predict whether a customer will churn or not based on their gender, education level, or similar.